
Course Project 1
=====================================

# download file 
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
```

# unzip data and read
```{r}
unzip("activity.zip")
stepdata <- read.csv("activity.csv", header = TRUE)
head(stepdata)
```

# Total number of steps by day
```{r}
library(magrittr)
library(dplyr)
databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps)) %>%na.omit()
hist(databydate$tsteps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
```

# Mean and median of steps per day
```{r}
mean(databydate$tsteps)
median(databydate$tsteps)
```

# Time series plot
```{r}
library(ggplot2)
databyinterval <- stepdata%>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps)) 
ggplot(databyinterval, aes(x=interval, y=tsteps))+ geom_line()
```

#The 5-minute interval that, on average, contains the maximum number of steps
```{r}
databyinterval[which(databyinterval$tsteps== max(databyinterval$tsteps)),]
```

#Code to describe and show a strategy for imputing missing data
```{r}
missingVals <- sum(is.na(data))
missingVals
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```

#Histogram of the total number of steps taken each day after missing values are imputed
```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
summary(FullSummedDataByDay)
hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```


#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )

meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")

ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
  facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
  ggtitle("Comparison of Average Number of Steps in Each Interval")
```


